---
title: "Final-Project"
author: "Mujaheed Abdulmalik"
date: "2023-07-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  DATASET  

```{r}
data_2 <- readxl::read_xlsx("Maths.csv") #import dataset

```


# LIBRARIES
```{r}
library(rstatix) #library for statstical tools
library(ez)
library(wPerm) #library for statstical tools like for non parameric one way anova
```



# IND ANOVA - PARENTAL RELATIONSHIP STATUS & STUDENT FINAL GRADE 

```{r}
#we have to group our data and get some descriptive statistics
data_pstatus <- data.frame(p_status = data_2$Pstatus, final_grade = data_2$G3)
parent.group <- group_by(data_pstatus, p_status )
get_summary_stats(parent.group, type = "mean_sd") #check for mean and standard deviation of both factors 
```

## ASSUMPTIONS  

Identify outliers 
```{r}
identify_outliers(parent.group, final_grade)

```


Second, we need to test the normality assumption.  

```{r}
shapiro_test(parent.group, final_grade)
```

```{r}
mod_1 <-  lm(final_grade ~ p_status, data = parent.group) 
par(mar = c(4,4,1.5, 1), mgp = c(2.5, 1, 0), mfrow =c(2,2))
plot(mod_1, which =1:3) #using the residual plots to check for normality
boxplot(final_grade ~ p_status, data = parent.group) #using boxplot to check for equal variance assumption
```

THe assumptions of equal variance was failed, a non parametric test using wperm will be used 

## TEST STATISTIC AND P VALUE  

```{r}
perm.oneway.anova(data_pstatus$final_grade, data_pstatus$p_status, R=3000)

```

## CONCLUSION  
 P-value is greater than alpha so we fail to reject null. 
 
 

\newpage


# IND ONE & TWO WAY ANOVA- MOTHER, FATHER OCCUPATION & STUDENT FINAL GRADE  


```{r}
#we have to group our data and get some descriptive statistics
data_Mjob <- data.frame(m_job = data_2$Mjob, final_grade = data_2$G3)
data_2$Mjob <- as.factor(data_2$Mjob)
m.group <- group_by(data_Mjob, m_job )
get_summary_stats(m.group, type = "mean_sd")
```

## ASSUMPTIONS  

Identify outliers 
```{r}
identify_outliers(m.group, final_grade)

```

 
```{r}
#check for assumptions using residuals plot
mod_2 <-  lm(final_grade ~ m_job, data = m.group) 
par(mar = c(4,4,1.5, 1), mgp = c(2.5, 1, 0), mfrow =c(2,2))
plot(mod_2, which =1:3)
boxplot(final_grade ~ m_job, data = m.group)
```
 
## TEST STAIISTIC AND P-VALUE   

```{r}
aov_2 <- aov(final_grade ~ m_job, data = m.group) #one way
summary(aov_2)
aov_two_way <- aov(G3 ~ Mjob * Fjob, data = data_2) #two-way
summary(aov_two_way)
```
 
 \
```{r}
#using tukeyhsd to check for the proportional difference in groups and account for adj p-value
TukeyHSD(aov(lm(final_grade ~ m_job, data = m.group)))  
``` 

## Effect size  

Check for effect size using eta squared  

```{r}
eta_squared(aov_2)
```


## Conclusion 

it is statistically significant, children moms who work under health tends to do bether than children with moms under others 

\newpage  

# IND ANOVA - FATHER OCCUPATION & STUDENT FINAL GRADE  


```{r}
#we have to group our data and get some descriptive statistics
data_Fjob <- data.frame(f_job = data_2$Fjob, final_grade = data_2$G3)
f.group <- group_by(data_Fjob, f_job )
get_summary_stats(f.group, type = "mean_sd")
```

## ASSUMPTIONS  

Identify outliers 
```{r}
identify_outliers(f.group, final_grade)

```

 
```{r}
mod_f <-  lm(final_grade ~ f_job, data = f.group) 
par(mar = c(4,4,1.5, 1), mgp = c(2.5, 1, 0), mfrow =c(2,2))
plot(mod_f, which =1:3)
boxplot(final_grade ~ f_job, data = f.group)
```
 
## TEST STAIISTIC AND P-VALUE   

```{r}
#use permutation as assumptions was violated
perm.oneway.anova(data_pstatus$final_grade, data_Fjob$f_job, R=3000) 
```

\newpage  


# IND ANOVA - FAMILY SIZE & STUDENT FINAL GRADE  


```{r}
#we have to group our data and get some descriptive statistics
data_Fsize <- data.frame(f_size = data_2$famsize, final_grade = data_2$G3)
size.group <- group_by(data_Fsize, f_size)
get_summary_stats(size.group, type = "mean_sd")
```

## ASSUMPTIONS  

Identify outliers 
```{r}
identify_outliers(size.group, final_grade)

```

 
```{r}
#plot for residuals to check assumptions 
mod_3 <-  lm(final_grade ~ f_size, data = size.group) 
par(mar = c(4,4,1.5, 1), mgp = c(2.5, 1, 0), mfrow =c(2,2))
plot(mod_3, which =1:3)
boxplot(final_grade ~ f_size, data = size.group)

```


## TEST STATISTIC AND P-VALUE  

```{r}
aov_3 <- aov(final_grade ~ f_size, data = size.group)
summary(aov_3)
```  




\newpage  




\newpage  

# IF MOTHERS EDUCATIONAL LEVEL DETERMINES THE STUDENT SUPPORT  

Contingency table  

<!-- pre processing data  -->
```{r}
table(data_2$Medu[data_2$famsup == "yes"])
table(data_2$Medu[data_2$famsup == "no"])
```
```{r}
#create a contingency table 
#list of yes and no values across groups
yes <- c(1,27,58,63,93)   
no <- c(1,33,45,36,39)

#fedu.df <- matrix(data = c(none, primary, grade_school, secondary, higher), nrow = 5, ncol = 2, byrow = TRUE)
MSupport <- matrix(data = c(yes, no), nrow = 2, ncol = 5, byrow = TRUE)
dimnames(MSupport) <- list(Support = c("Yes", "No"),
Education = c("none", "primary", "grade school", "secondary", "higher"))
MSupport
```


```{r}
#simulate chisq because assumptions werent met 
chisq_results <- chisq.test(x =  MSupport, correct = FALSE, simulate.p.value = T)
chisq_results
```
```{r}
chisq.posthoc.test::chisq.posthoc.test(MSupport) #Check effects and group differences
```

check effect size using cramers V  

```{r}
effectsize::cramers_v(MSupport) #check effect size using cramers V
```
## CONCLUSION  
it is statistically significant


\newpage  

# IF FATHERS EDUCATIONAL LEVEL DETERMINES THE STUDENT SUPPORT  

Contingency table  

<!-- pre processing data  -->
```{r}
table(data_2$Fedu[data_2$famsup == "yes"])
table(data_2$Fedu[data_2$famsup == "no"])
```


```{r}
#create a contingency table
#list of yes and no values across groups
yes <- c(1,39,67,64,71)   
no <- c(1,44,48,36,26)

#fedu.df <- matrix(data = c(none, primary, grade_school, secondary, higher), nrow = 5, ncol = 2, byrow = TRUE)
Support <- matrix(data = c(yes, no), nrow = 2, ncol = 5, byrow = TRUE)
dimnames(Support) <- list(Support = c("Yes", "No"),
Education = c("none", "primary", "grade school", "secondary", "higher"))
Support

```



## Test statisitc and p value   

```{r}
#simulate chisquare as assumptions werent met
chisq_results <- chisq.test(x =  Support, correct = FALSE, simulate.p.value = T) 
chisq_results
```
```{r}
chisq.posthoc.test::chisq.posthoc.test(Support) #conduct a posthoc test for proprotional differences 
```
```{r}

# store observed and expected
observed_values <- chisq_results$observed
expected_values <- chisq_results$expected

# Create a bar chart
barplot(rbind(observed_values, expected_values), beside = TRUE, col = c("blue", "red", "yellow", "green"), ylab = "Frequency", main = "Observed vs Expected Values For Fathers Education across Support")
legend("topleft", legend = c("Observed Yes", "Observed No", "Expected Yes", "Expected No"), fill = c("blue", "red", "yellow", "green"))
```

## CONCLUSION  
it is statistcally significant  


## Effect size
```{r}
effectsize::cramers_v(Support) #check effect size using cramers V
```




